---
title: "Flux Calculation with QEA Method"
output: html_document
author: Anas Emad
date: 2023-12-27
---

This document outlines a simple calculation of the fluxes using the quantized
eddy accumulation (QEA) method. We generate some synthetic data and simulate the
applicaiton of the QEA method then calculate the fluxes

```{r setup, include = FALSE}
# We define the quantization function as described in the accompanying paper
quantize_w <- function(w, 
                       quantize_threshold, 
                       w_full_scale_value) {
    residual_error_updraft   <- 0
    residual_error_downdraft <- 0
    N <- length(w)
    w_qua <- rep(0, N)

    for (i in seq_along(w)) {
        # Error diffusion is carried out within the respective wind direction
        if (w[i] > 0) {
            # Add error from previous step to modified w 
            w_mod <- w[i] - residual_error_updraft 
            # Quantize modified wind
            w_qua[i] <- ifelse(abs(w_mod) > quantize_threshold, 
                               sign(w_mod)*w_full_scale_value, 0)
            # Calculate residual error
            residual_error_updraft <- w_qua[i] - w_mod
        } else {
            w_mod <- w[i] - residual_error_downdraft
            w_qua[i] <- ifelse(abs(w_mod) > quantize_threshold, 
                               sign(w_mod)*w_full_scale_value, 0)
            residual_error_downdraft <- w_qua[i] - w_mod
        }

    }
    return(w_qua)
}

# We define the biased covariance function

bcov <- function (x,y) {
    N <- length(x)
    cov(x,y, use = "na.or.complete")*(N-1)/N
}
```


## Data Generation

We first generate some synthetic data for variables `w` (wind) and `C`
(scalar concentration). 
The generated time series is equivalent to 30 minutes of 10 Hz sampled data.


```{r data-generation}
N  <- 18000

set.seed(17)
ar_params  <- list(ar = c(0.993499, 0.017448, -0.218639, 0.136650))

w <- arima.sim(n = N, ar_params, sd = 2)

set.seed(18)
C <- arima.sim(n = N, ar_params, sd = 2.5) - w * 0.6 + 400

# We consider a nonzero mean of vertical wind in the range of 0.02 m
w <- w - mean(w) + 0.02

plot(w, C, pch = 20, col = rgb(0, 0, 0, 0.3), cex = 0.6, 
     xlab = expression("w (m s"^-1*")"), 
     ylab = expression("C (" * mu * "mol m"^-3*")"))
```

## Quantization of wind

Next, we quantize the wind data and calculate the quantization error.

```{r quantization}
quantization_threshold <- sd(w) * 2
w_full_scale_value     <- sd(w) * 4.4

wq                     <- quantize_w(w, quantization_threshold, w_full_scale_value)
wqr                    <- w - wq
```


# Flux Calculation

## Quantities Needed for Flux Calculation

Below are the quantities required for calculating the flux


```{r flux-quantities}
Cq_up_mean   <- mean(C[wq>0])   # Mean concentration in updraft reservoir (umol m^-3)
Cq_down_mean <- mean(C[wq<0])   # Mean concentration in downdraft reservoir
wq_up_mean   <- mean(wq[wq>0])  # Mean wq when wq > 0 which equals w_full_scale_value (m s^-1)
wq_down_mean <- mean(wq[wq<0])  # Mean wq when wq < 0 which equals -w_full_scale_value
N            <- length(w)       # Total count of w samples
Nq           <- sum(wq!=0)      # Count of collected samples
Nq_up        <- sum(wq>0)       # Count of collected samples in updraft
Nq_down      <- sum(wq<0)       # Count of collected samples in downdraft
alpha_co2_q  <- bcov(C, abs(wq))/bcov(C, wq) # Alpha: atmospheric asymmetric coefficient 

```

The estimation of alpha and its value is discussed in:
[(Emad and Siebicke, 2023)](https://doi.org/10.5194/amt-16-29-2023).

Next, we calculate the flux and the quantization flux error,
following the equations provided in the paper.

```{r flux-calculation}
# First order approximation of the flux (Eq. 20 in the paper)
# Unit of the flux ( umol m^-2 s^-1 )
Flux1 <- Cq_up_mean*Nq_up/N*wq_up_mean + 
         Cq_down_mean*Nq_down/N*wq_down_mean - 
         (Cq_up_mean*Nq_up/Nq + Cq_down_mean*Nq_down/Nq)*mean(wq)

# Quantization flux error (covariance of the scalar with the quantization error)
Quantization_flux_error <- bcov(C, wqr)

# Error ratio of the flux
Quantization_flux_error/Flux1

# Nonzero w correction (Eq 21 in the paper)
Nonzero_w_correction <- 1 / (1 - alpha_co2_q * mean(wq)/mean(abs(wq))) 
```
We find the relative error due to quantization is lower than 1% of the flux.

## Verification

We can verify the covariance is equal to the flux calculated using the QEA
method and check the quantized flux only.

```{r verification}
# Verify covariance equals flux calculated with QEA
all.equal(bcov(C, w), Flux1 * Nonzero_w_correction + Quantization_flux_error)

# Verify quantized flux only
all.equal(bcov(C, wq), Flux1 * Nonzero_w_correction)
```

